---
title: "Assignment-3"
author: "Bidit Sadhukhan"
date: '`r Sys.Date()`'
output: 
  html_document: 
    toc: yes
    number_sections: yes
    theme: readable
    highlight: zenburn
editor_options: 
  markdown: 
    wrap: 72
---

1.  Find the number of rows having NA month-wise in the airquality
    dataframe using (i) by (ii) aggregate (iii) table (iv) dplyr.

```{r}
dat=airquality
 head(dat)
 f4 <- function(x)
 {
   sum(is.na(x))
 }
#i)
 with(dat,by(dat,Month,function(x){sapply(x,f4)})) 
#ii)
    
 aggregate(dat,by=list(dat$Month),f4)
#iii)
 with(dat,table(complete.cases(dat),Month))
#iv)
 library(dplyr)
airquality %>%
  group_by(Month) %>%
  summarise_all(f4)
```

2.  Recode Temp variable: create a new col with values of Temp into
    various categories such as Extreme, High, Medium, Low depending upon
    Temp value using (i) dplyr and ifelse (ii) dplyr and a custom
    function (hint:use sapply to call the function) (iii) cut function
    (iv) simple indexing of rows by conditions (use of index vector
    mechanism) (v) replace function. Verify if the results match using
    table function.

```{r}
#i)
dat = dat %>%
  mutate(Temp_new=ifelse(Temp>quantile(Temp,.75),"Extreme",ifelse(Temp>quantile(Temp,.5),"High",ifelse(Temp>quantile(Temp,.25),"Medium","Low"))),.after=Temp) 

head(dat)
  
#ii)
f1 <- function(x){
  
}
table(sapply(dat$Temp,f1))
dat$Temp
#iii)
brks <- quantile(dat$Temp,probs = seq(0,1,0.25)); brks
dat$Temp_new2<- cut(dat$Temp,breaks = brks, labels = c("Low","Medium","High","Extreme"),include.lowest = T)
head(dat)

#iv)
 #dat$Temp[which(dat$temp_3>quantile(dat$Temp,0.75))]
 
# dat[dat$temp_3 > quantile(dat$temp_3,0.75),]="Extreme"
 #dat$Temp

#v)
dat$Temp_lvl3=replace(dat$Temp,dat$Temp>quantile(dat$Temp,.75),"Extreme")

dat$Temp_lvl3=replace(dat$Temp_lvl3,dat$Temp>quantile(dat$Temp,.5)&dat$Temp<=quantile(dat$Temp,.75),"High")

dat$Temp_lvl3=replace(dat$Temp_lvl3,dat$Temp>quantile(dat$Temp,.25)&dat$Temp<=quantile(dat$Temp,.5),"Medium")

dat$Temp_lvl3=replace(dat$Temp_lvl3,dat$Temp<=quantile(dat$Temp,.25),"Low")
head(dat)
```

3.  Replace the NA values in Ozone of air quality dataframe with the
    mean value of Ozone column using dplyr.

```{r}
library(dplyr)
dat %>% group_by(Month) %>% mutate_at(vars(Ozone), ~ifelse(is.na(.), mean(., na.rm = TRUE), .)) 

#without using dplyr
dat$Ozone[is.na(dat$Ozone)]=mean(dat$Ozone,na.rm=T)
head(dat)

```

4.  Replace the NA values in Ozone with the mean value of Ozone column
    in airquality dataframe where the mean value of Ozone for the same
    month is applied using (i) dplyr (ii) subset (iii) aggregate
    function and ifelse (iv) aggregate and mapply (iv) tapply and custom
    function.

```{r}
library(dplyr)
dat %>% 
  group_by(Month) %>% 
  mutate_at(vars(Ozone), ~ifelse(is.na(.), mean(., na.rm = TRUE), .)) 
  
```

5.  How to reorder a dataframe by particular column(s) w/o using dplyr -
    order the air quality dataframe by the Month and Temp columns
    together w/o using dplyr.

```{r}
dat_1 <- airquality
head(dat_1)
head(arrange(dat_1,Month,Temp))
```

6.  Using a mpg dataset create the following graph

```{r}
if(!("ggplot2" %in% installed.packages())) install.packages("ggplot2")
library(ggplot2)
ggplot(mpg,aes())+
  geom_bar(mapping=aes(x=manufacturer,fill=manufacturer))

```

7.  Modify the above plot to include the following data (data on the
    models for each manufacturer):

```{r}
ggplot(mpg,aes())+
  geom_bar(mapping=aes(x=manufacturer,fill=model))+coord_flip()

```

8.  Try to modify the plot in q.no. 1 to reorder the bars in the order
    of the count. (Hint: use reorder function in the y value of the aes
    function)

```{r}
ggplot(mpg,aes())+
  geom_bar(mapping=aes(x=reorder(manufacturer,manufacturer,NROW),fill=manufacturer,y=..count..))+coord_flip()+xlab("Manufacturer_ordered")+ylab("Count")+coord_flip()

```

9.  Build a plot of the count of cars categorised by model and
    manufacturer. Align the x axis labels vertically so that they are
    visible as shown below (Hint: use theme function along with
    element_text function to set the value of axis.text.x)

```{r}
library(ggplot2)
ggplot(mpg,aes())+
  geom_count(aes(x=manufacturer,y=model,colour=manufacturer))+theme(axis.text.x  =element_text(colour = "blue",angle = 90) )

```

10. Find the number of records for each combination of manufacturer and
    model (i) using dplyr (ii) not using dplyr (iii) using the melt
    function in the reshape package. This question basically finds the
    underlying data for the plot in q.no. 4.

```{r}
if(!("reshape" %in% installed.packages())) install.packages("reshape")
#(i)
mpg %>% group_by(manufacturer, model) %>% summarise(n())

#(ii)
with(mpg,table(manufacturer,model))

#iii)

```

11. Annotate the plot in q.no. 4 with text labels indicating the largest
    group (largest circle). (Hint: use the annotate function).

```{r}

```

12. Plot the columns Month and Day along with Temp from the dataset
    airquality in a heat map (raster plot) as below. The colours in the
    plot should range between yellow and red as shown below (Hint: use
    scale_fill_gradient)

```{r}
ggplot(data=airquality,aes(x=Month,y=Day))+
geom_raster(aes(fill=Temp))+
  scale_fill_gradient(high="red",low="yellow")  
```





13. Experiment with violin plot on the movies dataset (from ggplot2
    package) using the variables rating (x axis) and votes (y axis). Try
    different scales (default, scale_y\_log10, reciprocal, sqrt,
    manually setting y limit values using ylim function with a limit
    between say 0 and 30). You may also experiment with
    scale_y\_continuous function example:

```{r}
library("ggplot2movies")
m <- movies;head(m)
ggplot(data = m) +geom_violin(mapping=aes(rating,votes, fill= cut_width(rating,0.5)),scale = "area")
ggplot(data = m) +geom_violin(mapping=aes(rating,votes, fill= cut_width(rating,0.5)))+ scale_y_log10()
ggplot(data = m) +geom_violin(mapping=aes(rating,votes, fill= cut_width(rating,0.5)))+ scale_y_continuous(trans = "reciprocal")
ggplot(data = m) +geom_violin(mapping=aes(rating,votes, fill= cut_width(rating,0.5)))+ scale_y_sqrt()  

ggplot(data = m) +geom_boxplot(mapping=aes(rating,votes, fill= cut_width(rating,0.5)),scale = "area")
ggplot(data = m) +geom_boxplot(mapping=aes(rating,votes, fill= cut_width(rating,0.5)))+ scale_y_log10()
ggplot(data = m) +geom_boxplot(mapping=aes(rating,votes,fill= cut_width(rating,0.5)))+ scale_y_continuous(trans = "reciprocal")
ggplot(data = m) +geom_boxplot(mapping=aes(rating,votes, fill= cut_width(rating,0.5)))+ scale_y_sqrt()  
```




14. Attempt to plot an area plot (geom_area) with the column votes in
    movies dataset. Notice the extreme skewness of the data. Experiment
    with various ranges for the votes column in order to get a
    meaningful plot (say xlim between 0 and 200). Also attempt to use a
    transformed value of x (votes) using log10, i.e. using
    scale_x\_continuous(trans="log10").

```{r}
library("ggplot2movies")
m <- movies;head(m)
ggplot(data=m, aes(votes)) + geom_area(stat = "bin")
ggplot(data=m, aes(votes)) + geom_area(stat = "bin") + xlim(0,200)
ggplot(data=m, aes(votes)) + geom_area(stat = "bin") + scale_x_continuous(trans = "log10")

```


15. Attempt a scatter plot of mpg dataset using geom_jitter and the cty and hwy columns. Modify
the plot by customising your x axis labels and x axis ticks and shape based on a categorical
columns. Customise the legend using the theme function.

```{r}
ggplot(mpg,aes())+
  geom_jitter(aes(x=cty,y=hwy)) 
ggplot(mpg,aes())+
  geom_jitter(aes(x=cty,y=hwy, shape = fl)) +
  labs(x="cty_2") + 
  theme_linedraw()

```