---
title: "R Slides PDS"
author: "Bidit Sadhukhan"
date: "`r Sys.Date()`"
output: 
  html_document: 
    theme: readable
    number_sections: yes
    toc: yes
    highlight: zenburn
    df_print: kable
---


# Slide-3 
## R data structures and functions

1. Given the following vectors use mapply to find the sum of all the first elements of the numerical vectors, sum of all the second elements and so on
+ v1 <- c(1,4,9,1)

+ v2 <- c('a','b','b','a')

+ v3 <- c(2,3,4,1)

+ v4 <- c(3,1,3,1)
```{r}
v1 <- c(1,4,9,1);v1
v2 <- c("a","b","b","a");v2
v3 <- c(2,3,4,1);v3
v4 <- c(3,1,3,1);v4
mapply(sum,v1,v3,v4)
```
2. Create 3 vectors of the same length containing information about the age, marks, number of books read in a year and the gender for 5 students in a class.
Combine these vectors into a dataframe. Use tapply function to find the average number of marks obtained for each gender group and the min number of books
read for each gender group.
```{r}
set.seed(2569)
age <- sample(10:15,5);age
marks <- round(runif(5,25,100),1);marks
no_books <- sample(0:10,5);no_books
gender <- sample(c("M","F"),5, replace = T);gender
df <-  data.frame(age,marks,no_books,gender);df
tapply(df$marks,df$gender, mean)
tapply(df$no_books,df$gender,min)
```

3. With the above dataframe created use the apply function to find the average marks obtained by the students, average number of books read and the average
age of the students.
```{r}
apply(df[,1:3],2,mean)
```
4. Create a matrix using the 3 numerical vectors and use the apply function to find the average marks obtained by the students, average number of books read and
the average age of the students.
```{r}
m <- matrix(c(age,marks,no_books),nrow=5,dimnames = list(1:5,c("age","marks","no.books")));m
round(apply(m,2,mean))
```
5. With the above dataframe created use the lapply function to find the average marks obtained by the students, average number of books read and the average
age of the students.
```{r}
lapply(df[,1:3],mean)
```
6. Create 2 vectors of the same length containing the before training and after training measurement of competency on a scale of 1-10 for 5 students. For example,
let vector 1 contains the pre-training competency score of the 5 students as 2,7,4,9,6 and the post training competency score as 5,6,7,6,6 respectively. Use
mapply on these 2 vectors to find the difference between post and pre scores. Hint: use minus operator with quotes as a supplied function, i.e. mapply('-',...)
```{r}
set.seed(2589)
v1 <- sample(1:10,5);v1
v2 <- sample(1:10,5);v2
mapply('-',v1,v2)
```

### Session Info
```{r}
sessionInfo()
```
