---
title: "R Slides PDS"
author: "Bidit Sadhukhan"
date: "`r Sys.Date()`"
output: 
  html_document: 
    theme: readable
    number_sections: yes
    toc: yes
    highlight: zenburn
    df_print: kable
---


# Slide-3 
## R data structures and functions

1. Given the following vectors use mapply to find the sum of all the first elements of the numerical vectors, sum of all the second elements and so on
+ v1 <- c(1,4,9,1)

+ v2 <- c('a','b','b','a')

+ v3 <- c(2,3,4,1)

+ v4 <- c(3,1,3,1)
```{r}
v1 <- c(1,4,9,1);v1
v2 <- c("a","b","b","a");v2
v3 <- c(2,3,4,1);v3
v4 <- c(3,1,3,1);v4
mapply(sum,v1,v3,v4)
```
2. Create 3 vectors of the same length containing information about the age, marks, number of books read in a year and the gender for 5 students in a class.
Combine these vectors into a dataframe. Use tapply function to find the average number of marks obtained for each gender group and the min number of books
read for each gender group.
```{r}
set.seed(2569)
age <- sample(10:15,5);age
marks <- round(runif(5,25,100),1);marks
no_books <- sample(0:10,5);no_books
gender <- sample(c("M","F"),5, replace = T);gender
df <-  data.frame(age,marks,no_books,gender);df
tapply(df$marks,df$gender, mean)
tapply(df$no_books,df$gender,min)
```

3. With the above dataframe created use the apply function to find the average marks obtained by the students, average number of books read and the average
age of the students.
```{r}
apply(df[,1:3],2,mean)
```
4. Create a matrix using the 3 numerical vectors and use the apply function to find the average marks obtained by the students, average number of books read and
the average age of the students.
```{r}
m <- matrix(c(age,marks,no_books),nrow=5,dimnames = list(1:5,c("age","marks","no.books")));m
round(apply(m,2,mean))
```
5. With the above dataframe created use the lapply function to find the average marks obtained by the students, average number of books read and the average
age of the students.
```{r}
lapply(df[,1:3],mean)
```
6. Create 2 vectors of the same length containing the before training and after training measurement of competency on a scale of 1-10 for 5 students. For example,
let vector 1 contains the pre-training competency score of the 5 students as 2,7,4,9,6 and the post training competency score as 5,6,7,6,6 respectively. Use
mapply on these 2 vectors to find the difference between post and pre scores. Hint: use minus operator with quotes as a supplied function, i.e. mapply('-',...)
```{r}
set.seed(2589)
v1 <- sample(1:10,5);v1
v2 <- sample(1:10,5);v2
mapply('-',v1,v2)
```


# Assignment-2
1. Specieswise tabulate number of missing values for the column root of the data frame rainforest (DAAG) Hint: use complete.cases
```{r}
library(DAAG)
df3 <- rainforest
library(dplyr)
with(df3,aggregate(root ~ species,df3,function(root){sum(is.na(root))},na.action = NULL))

```

2. For each column of the data frame Pima.tr2 (MASS ), determine the number of missing values, broken down by type.
(a) with sapply, and subset command. 
(b) with by
(c) with group by 
(d) with aggregate

```{r}
library(MASS)
df4 <-  Pima.tr2
search()   # to see the order of package as there was a statement in R attaching hill from DAAG package
f4 <- function(x)
{
  sum(is.na(x))
}
#a)
for (lev in levels(df4$type)) {
  print(sapply(subset(df4,type==lev),f4))
}

#b)
with(df4,by(df4,type,function(x){sapply(x,f4)}))

#c)library(dplyr)
df4 %>%
  group_by(type) %>%
  summarise(missing_npreg=f4(npreg),
            missing_glu=f4(glu),
            missing_bp=f4(bp),
            missing_skin=f4(skin),
            missing_bmi = f4(bmi),
            missing_ped = f4(ped),
            missing_age = f4(age)
            )

#d)
with(df4,aggregate(df4,by=list(type),f4))
```
3. In the data frame rainforest (package - DAAG), use table(rainforest$species) to check the names and numbers of each species that are present in the data.
```{r}
df3 <- rainforest
table(rainforest$species)
```

4. Extract the following subsets from the data frame ais (DAAG): 
(a) data for the rowers
(b) data for the rowers, the netballers and the tennis players 
(c) data for the female basketabllers and rowers

```{r}
library(DAAG)
df <- ais
subset(df,sport == "Row")
subset(df,sport %in% c("Row","Netball","Tennis"))
subset(df,sport=="B_Ball" & sex=="f")
```

5. Using the ais (DAAG) dataframe do the following:
(a) recode wt column into a new categorical column such that there are equal (or almost equal) number of data in 10 equal sized intervals, name the intervals using a,b,c,...,etc. For example the result of applying the table function to the new column would given something as below:
 a  b  c  d  e  f  g  h  i  j 
20 20 21 19 22 18 21 19 21 20
(Hint: use a combination of quantile and cut functions. Use quantile function to derive the breakpoints and then apply it to the cut function with appropriate labels to derive the new column)
(b) create a new column with values given by the following condition:
bmi more than 22.96 and lbm more than or equal to 63 and sport is NOT any of the following: B_Ball, Field, Gym then then the new column value is A otherwise B
(Hint: check the use of %in% operator or match function and their inverse using ! operator)
      (i) do the part (b) using mapply and named functions
      (ii) do the part (b) w/o using mapply
      
```{r}
library(DAAG)
#a)
df <- ais
mini <- quantile(df$wt,probs = 0)
maxi <- quantile(df$wt,probs = 1)
tab_breaks <- with(df, seq(mini,maxi,(maxi-mini)/10))

df$factor_wt <- with(df, cut(wt,tab_breaks,labels=LETTERS[1:10],include.lowest=TRUE))
head(df)

#b)i)
f3 <- function(x,y,z)
{
  if(x>22.6 & y>= 63 & !(z %in% c("B_Ball","Field","Gym")))
  {
    changed <- "A"
  }
    else{
      changed <- "B"
    }
  
}
df$changed <- as.factor(mapply(f3,df$bmi,df$lbm,df$sport))
str(df)


#b) ii)
library(dplyr)
 
 df %>%
   filter(bmi>22.6 & lbm >= 63 ) %>%
  filter(!(sport %in% c("B_Ball","Field","Gym")))
```

6. Using the dataframe ais (DAAG) create a new column using the sport column so that the "_" in the names is removed, i.e. B_Ball becomes BBall, etc. 
(Hint: the sport column is a factor column and not a character column; use combination of strsplit, unlist and paste function if required or any other method of your choice)

```{r}
library(DAAG)
df <- ais
y <- levels(df$sport);y
y[c(1,7,8,10)]=c("BBall","T400m","TSprnt","WPolo")
levels(df$sport) <- y
levels(df$sport)
```


### Session Info
```{r}
sessionInfo()
```

